### 2022-01-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@cf723b26cf48b6350d857f651e5d72035bb526cd 🚀

\begin{enumerate}
    \item Despite their importance in many ecological processes,
collecting data and information on ecological interactions is an
exceedingly challenging task. For this reason, large parts of the world
have a data deficit when it comes to species interactions, and how the
resulting networks are structured. As data collection alone is unlikely
to be sufficient, community ecologists must adopt predictive methods.%
    \item We present a methodological framework that uses graph
embedding and transfer learning to assemble a predicted list of trophic
interactions of a species pool for which their interactions are unknown.
Specifically, we `learn' the information from a known interaction
network by inferring the latent traits of species and infer the latent
traits of a species pool for which we have no \emph{a priori}
interaction data based on their phylogenetic relatedness to species from
the known network. The latent traits can then be used to predict
interactions and construct and interaction network.%
    \item Here we assembled a metaweb for Canadian mammals derived from
interactions in the European food web, despite only 4\% of common
species being shared between the two locations. The results of the
predictive model are compared against databases of recorded pairwise
interactions, showing that we correctly recover 91\% of known
interactions.%
    \item The framework itself is robust even when the known network is
incomplete or contains spurious interactions making it an ideal
candidate as a tool for filling gaps when it comes to species
interactions. We provide guidance on how this frameworl can be adapted
by substituting some approaches or predictors in order to make it more
generally applicable.%
\end{enumerate}

@@

Here, we present a general method to recommend potential trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang, 2010;
Torrey \& Shavlik, 2010). In this instance, we solve the problem of
predicting trophic interactions between species, based on knowledge
extracted from another species pool for which interactions are known by
using phylogenetic structure as a medium for transfer. There is a
plurality of measures of species similarities that can be used for
inferring \emph{potential} species interactions \emph{i.e.} metaweb
reconstruction (see \emph{e.g.} Morales-Castilla et al., 2015); however,
phylogenetic proximity has several desirable properties when working at
large scales. Gerhold et al. (2015) made the point that phylogenetic
signal captures diversification of characters (large macro-evolutionary
process), but not necessarily community assembly (fine ecological
process); Dormann et al. (2010) previously found very similar
conclusions. Interactions tend to reflect a phylogenetic signal because
they have a conserved pattern of evolutionary convergence that
encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares et al., 2009; Mouquet et al., 2012), and - most
importantly - retain this signal even when it is not detectable at the
community scale (Hutchinson et al., 2017; Poisot \& Stouffer, 2018).
Finally, species interactions at macro-ecological scales seem to respond
mostly to macro-evolutionary processes (Price, 2003); which is evidenced
by the presence of conserved backbones in food webs (Dalla Riva \&
Stouffer, 2016; Mora et al., 2018), strong evolutionary signature on
prey choice (Stouffer et al., 2012), and strong phylogenetic signature
in food web intervality (Eklöf \& Stouffer, 2016). Phylogenetic
reconstruction has also previously been used within the context of
ecological networks, namely understanding ancestral plant-insect
interactions (Braga et al., 2021). Taken together, these considerations
suggest that phylogenies can reliably be used to transfer knowledge on
species interactions.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (latent vulnerability traits are more
representative of species at the lower trophic-level and latent
generality traits are more representative of species at higher
trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (known interactions; Maiorano et al.,
2020a, 2020b) and, based on phylogenetic relationships between mammals
globally (\emph{i.e.}, phylogenetic tree Upham et al., 2019), infer a
metaweb for the Canadian mammalian species pool (interactions are
treated as unknown in this instance). Following the definition of Dunne
(2006), a metaweb is a network analogue to the regional species pool;
specifically, it is an inventory of all \emph{potential} interactions
between species from a spatially delimited area (and so captures the
\(\gamma\) diversity of interactions). The metaweb is, therefore,
\emph{not} a prediction of the food web at a specific locale within the
spatial area it covers, and will have a different structure (notably by
having a larger connectance; see \emph{e.g.} Wood et al., 2015). These
local food webs (which captures the \(\alpha\) diversity of
interactions) are a subset of the metaweb's species and interactions,
and have been called ``metaweb realizations'' (Poisot et al., 2015).
Differences between local food web and their metaweb are due to chance,
species abundance and co-occurrence, local environmental conditions, and
local distribution of functional traits, among others.

@@

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%. It should be reiterated that the framework presented
in fig.~\ref{fig:concept} is amenable to changes; notably, the measure
of similarity may not be phylogeny, and can be replaced by information
on foraging (Beckerman et al., 2006), cell-level mechanisms (Boeckaerts
et al., 2021), or a combination of traits and phylogenetic structure
(Stock, 2021). Most importantly, although we focus on a trophic system,
it is an established fact that different (non-trophic) interactions do
themselves interact with and influence the outcome of trophic
interactions (see \emph{e.g.} Kawatsu et al., 2021; Kéfi et al., 2012).
Future development of metaweb inference techniques should cover the
prediction of multiple interaction types.

@@

We use data from the European metaweb assembled by Maiorano et al.
(2020a). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
This metaweb itself is a network of binary (\emph{i.e.}
presence/absence), potential two-way interactions between species pairs.

We filtered down the European metaweb to create a subgraph corresponding
to all mammals by matching species names in the original network to the
Global Biodiversity Information Facility (GBIF) taxonomic backbone (GBIF
Secretariat, 2021) and retaining all those who matched to mammals. This
serves a dual purpose 1) to extract only mammals from the European
network and 2) to match and standardize species names when aggregating
the different data sources further downstream (which is an important
consideration when combining datasets (Grenié et al., 2021)). All nodes
had valid matches to GBIF at this step, and so this backbone is used for
all name reconciliation steps as outlined below.

The European metaweb represents the knowledge we want to learn and
transfer; the phylogenetic similarity of mammals here represents the
information for transfer (\emph{i.e.} the transfer medium). We used the
mammalian consensus supertree by Upham et al. (2019), for which all
approximatively 6000 names have been similarly matched to their GBIF
valid names. This step allows us to place each node of the mammalian
European metaweb in the phylogeny.

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the International Union for Conservation of Nature
(IUCN) checklist, and selected the terrestrial and semi-aquatic species
(this corresponds to the same selection that was applied by Maiorano et
al. (2020a) in the European metaweb). The IUCN names were, as
previously, reconciled against GBIF to have an exact match to the
taxonomy.

After taxonomic cleaning and reconciliation as outlined in the following
sections, the mammalian European metaweb has 260 species, and the
Canadian species pool has 163; of these, 17 (about 4\% of the total) are
shared, and 89 species from Canada (54\%) had at least one congeneric
species in Europe. The similarity for both species pools predictably
increases with higher taxonomic order, with 19\% of shared genera, 47\%
of shared families, and 75\% of shared orders; for the last point,
Canada and Europe each had a single unique order (\emph{Didelphimorphia}
for Canada, \emph{Erinaceomorpha} for Europe).

In the following sections, we describe the representational learning
step applied to European data, the transfer step through phylogenetic
similarity, and the generation of a probabilistic metaweb for the
destination species pool.

@@

The core point of our method is the transfer of knowledge of a known
ecological network, in order to predict interactions between species
from another location at which the network is unknown (or partially
known). In fig.~\ref{fig:concept}, we give a high-level overview of the
approach; in the example around which this manuscript is built
(leveraging detailed knowledge about binary trophic interactions between
Mammalia in Europe to predict the less known trophic interactions
between closely phylogenetically related Mammalia in Canada), we use a
series of specific steps for network embedding, trait inference, network
prediction and thresholding.

Specifically, our approach can be summarized as follows: from the known
network in Europe, we use a truncated Singular Value Decomposition
(t-SVD; Halko et al., 2011) to generate latent traits representing a
low-dimensional embedding of the network. As an aside, most ecologists
are indirectly familiar with SVD: Principal Component Analysis is a
special case of SVD, which is more sensitive to numerical instabilities
(see notably Shlens, 2014). The latent traits give an unbiased estimate
of the node's position in the latent feature spaces and can be mapped
onto a reference phylogeny (other distance-based measures of species
proximity that allow for the inference of features in the latent space
can be used, for example the dissimilarity in functional traits). Based
on the reconstructed latent traits for species in the destination
species pool, a Random Dot Product Graph model (hereafter RDPG; S. J.
Young \& Scheinerman, 2007) predicts the interaction between species
through a function of the nodes' features through matrix multiplication.
Thus, from latent traits and node position, we can infer interactions.

The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
(Holm, 2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information than we might think; as such, even algorithms that are
disconnected from the model can make correct guesses most of the time
(Halevy et al., 2009); in fact, in an instance of ecological forecasting
of spatio-temporal systems, model-free approaches (\emph{i.e.} drawing
all of their information from the data) outperformed model-informed ones
(Perretti et al., 2013).

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy et al., 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer, 2016); in other words, the latent variables of an RDPG
can be mapped onto a phylogenetic tree, and phylogenetically similar
predators should share phylogenetically similar preys. In addition,
recent advances show that the latent variables produced this way can be
used to predict \emph{de novo} network edges. Interestingly, the latent
variables do not need to be produced by decomposing the network itself;
in a recent contribution, Runghen et al. (2021) showed that deep
artificial neural networks are able to reconstruct the left and right
subspaces of an RDPG, in order to predict human movement networks from
individual/location metadata. This is an exciting opportunity, as it
opens up the possibility of using additional metadata as predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom, Dalla
Riva, et al., 2021) and to allow highly accurate prediction of the
interactions within a single network (Poisot, Ouellet, et al., 2021).
Under SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.

@@

An RDPG estimates the probability of observing interactions between
nodes (species) as a function of the nodes' latent variables, and is a
way to turn a SVD (which decompose one matrix into three) into two
matrices that can be multiplied to provide an approximation of the
network. The latent variables used for the RDPG, called the left and
right subspaces, are defined as
\(\mathscr{L} = \mathbf{U}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{V}'\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R} = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R} \approx \mathbf{A}\). Using a rank of 1 for the
t-SVD provides a first-order approximation of the network. One advantage
of using a RDPG rather than a SVD is that the number of components to
estimate decreases; notably, one does not have to estimate the singular
values of the SVD. Furthermore, the two subspaces can be directly
multiplied to yield a network.

@@

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
Maiorano et al. (2020a) and O'Connor et al. (2020), this seems like a
reasonable assumption, although one that does not hold for all metawebs.
We used the thresholding approach presented in Poisot, Ouellet, et al.
(2021), and picked a cutoff that maximized Youden's \(J\) statistic (a
measure of the informedness (trust) of predictions; Youden (1950)); the
resulting cutoff was 0.22, and gave an accuracy above 0.99. In Supp.
Mat. 1, we provide several lines of evidence that using the entire
network to estimate the threshold does not lead to overfitting; that
using a subset of species would yield the same threshold; that
decreasing the quality of the original data by adding of removing
interactions would minimally affect the predictive accuracy of RDPG
applied to the European metaweb; and that the networks reconstructed
from artificially modified data are reconstructed with the correct
ecological properties.

@@

Because the confidence intervals on the inferred trait space are
probably over-estimates, we decided to apply a thresholding step to the
interactions after the data inflation (fig.~\ref{fig:thresholds}).
Cirtwill \& Hambäck (2021) proposed a number of strategies to threshold
probabilistic networks. Their methods assume the underlying data to be
tag-based sequencing, which represents interactions as co-occurrences of
predator and prey within the same tags; this is conceptually identical
to our Bernoulli-trial based reconstruction of a probabilistic network.
We performed a full analysis of the effect of various cutoffs, and as
they either resulted in removing too few interactions, or removing
enough interactions that species started to be disconnected from the
network, we set this threshold for a probability equivalent to 0 to the
largest possible value that still allowed all species to have at least
one interaction with a non-zero probability. The need for this slight
deviation from the Cirtwill \& Hambäck (2021) method highlights the need
for additional development on network thresholding.

@@

In fig.~\ref{fig:thresholds}, we examine the effect of varying the
cutoff on \(P(i \rightarrow j)\) on the number of links, species, and
connectance. Determining a cutoff using the maximum curvature, or
central difference approximation of the second order partial derivative,
as suggested by \emph{e.g.} Cirtwill \& Hambäck (2021), results in
species being lost, or almost all links being kept. We therefore settled
on the value that allowed all species to remain with at least one
interaction. This result, in and of itself, suggests that additional
methodological developments for the thresholding of probabilistic
networks are required.

@@

One important aspect in which Europe and Canada differ (despite their
comparable bioclimatic conditions) is the degree of the legacy of human
impacts, which have been much longer in Europe. Nenzén et al. (2014)
showed that even at small scales (the Iberian peninsula), mammal food
webs retain the signal of both climate change and human activity, even
when this human activity was orders of magnitude less important than it
is now. Similarly, Yeakel et al. (2014) showed that changes in human
occupation over several centuries can lead to food web collapse.
Megafauna in particular seems to be very sensitive to human arrival
(Pires et al., 2015). In short, there is well-substantiated support for
the idea that human footprint affects more than the risk of species
extinction (Marco et al., 2018), and can lead to changes in interaction
structure. Yet, owing to the inherent plasticity of interactions, there
have been documented instances of food webs undergoing rapid
collapse/recovery cycles over short periods of time (Pedersen et al.,
2017). The embedding of a network, in a sense, embeds its
macro-evolutionary history, especially as RDPG captures ecological
signal (Dalla Riva \& Stouffer, 2016); at this point, it is important to
recall that a metaweb is intended as a catalogue of all potential
interactions, which should then be filtered (Morales-Castilla et al.,
2015). In practice (and in this instance) the reconstructed metaweb will
predict interactions that are plausible based on the species'
evolutionary history, however some interactions would/would not be
realized due to human impact.

@@

prioritized. Although RDPG was able to maintain very high predictive
power when interactions were missing, the addition of false positive

@@

\textbf{Authors' contributions:} TS, SB, and TP designed the study and
performed the analysis; GVDR, MF, and RR provided additional feedback on
the analyses. DC, BM, and FB helped with data collection. All editors
contributed to writing and editing the manuscript.

>>>>

\begin{enumerate}
    \item Despite their importance in many ecological processes,
collecting data and information on ecological interactions is an
exceedingly challenging task. For this reason, large parts of the world
have a data deficit when it comes to species interactions, and how the
resulting networks are structured. As data collection alone is unlikely
to be sufficient, community ecologists must adopt predictive methods.%
    \item We present a methodological framework that uses graph
embedding and transfer learning to assemble a predicted list of trophic
interactions of a species pool for which their interactions are unknown.
Specifically, we `learn' the information from a known interaction
network by inferring the latent traits of species and infer the latent
traits of a species pool for which we have no \emph{a priori}
interaction data based on their phylogenetic relatedness to species from
the known network. The latent traits can then be used to predict
interactions and construct an interaction network.%
    \item Here we assembled a metaweb for Canadian mammals derived from
interactions in the European food web, despite only 4\% of common
species being shared between the two locations. The results of the
predictive model are compared against databases of recorded pairwise
interactions, showing that we correctly recover 91\% of known
interactions.%
    \item The framework itself is robust even when the known network is
incomplete or contains spurious interactions making it an ideal
candidate as a tool for filling gaps when it comes to species
interactions. We provide guidance on how this framework can be adapted
by substituting some approaches or predictors in order to make it more
generally applicable.%
\end{enumerate}

@@

Here, we present a general method to infer potential trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang, 2010;
Torrey \& Shavlik, 2010). In this instance, we solve the problem of
predicting trophic interactions between species, based on knowledge
extracted from another species pool for which interactions are known by
using phylogenetic structure as a medium for transfer. There is a
plurality of measures of species similarities that can be used for
inferring \emph{potential} species interactions \emph{i.e.} metaweb
reconstruction (see \emph{e.g.} Morales-Castilla et al., 2015); however,
phylogenetic proximity has several desirable properties when working at
large scales. Gerhold et al. (2015) made the point that phylogenetic
signal captures diversification of characters (large macro-evolutionary
process), but not necessarily community assembly (fine ecological
process); Dormann et al. (2010) previously found very similar
conclusions. Interactions tend to reflect a phylogenetic signal because
they have a conserved pattern of evolutionary convergence that
encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares et al., 2009; Mouquet et al., 2012), and - most
importantly - retain this signal even if it is obscured at the community
scale due to \emph{e.g.} local conditions (Hutchinson et al., 2017;
Poisot \& Stouffer, 2018). Finally, species interactions at
macro-ecological scales seem to respond mostly to macro-evolutionary
processes (Price, 2003); which is evidenced by the presence of conserved
backbones in food webs (Dalla Riva \& Stouffer, 2016; Mora et al.,
2018), strong evolutionary signature on prey choice (Stouffer et al.,
2012), and strong phylogenetic signature in food web intervality (Eklöf
\& Stouffer, 2016). Phylogenetic reconstruction has also previously been
used within the context of ecological networks, namely understanding
ancestral plant-insect interactions (Braga et al., 2021). Taken
together, these considerations suggest that phylogenies can reliably be
used to transfer knowledge on species interactions.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{figures/figure-concept_v2.png}
\caption{Overview of the phylogenetic transfer learning (and prediction)
of species interactions networks. Starting from an initial, known,
network, we learn its representation through a graph embedding step
(here, a truncated Singular Value Decomposition; Step 1), yielding a
series of latent traits (latent vulnerability traits are more
representative of species at the lower trophic-level and latent
generality traits are more representative of species at higher
trophic-levels; \emph{sensu} Schoener (1989)); second, for the
destination species pool, we perform ancestral character estimation
using a phylogeny (here, using a Brownian model for the latent traits;
Step 2); we then sample from the reconstructed distribution of latent
traits (Step 3) to generate a probabilistic metaweb at the destination
(here, assuming a uniform distribution of traits), and threshold it to
yield the final list of interactions (Step 4).}\label{fig:concept}
}
\end{figure}

In fig.~\ref{fig:concept}, where we provide a methodological overview
based on learning the embedding of a metaweb of trophic interactions for
European mammals (known interactions; Maiorano et al., 2020a, 2020b)
and, based on phylogenetic relationships between mammals globally
(\emph{i.e.}, phylogenetic tree Upham et al., 2019), infer a metaweb for
the Canadian mammalian species pool (using only a species list
\emph{i.e.} interactions are `unknown' in this instance). Following the
definition of Dunne (2006), a metaweb is a network analogue to the
regional species pool; specifically, it is an inventory of all
\emph{potential} interactions between species from a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is, therefore, \emph{not} a prediction of the food web at a
specific locale within the spatial area it covers, and will have a
different structure (notably by having a larger connectance; see
\emph{e.g.} Wood et al., 2015). These local food webs (which captures
the \(\alpha\) diversity of interactions) are a subset of the metaweb's
species and interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local food web and their
metaweb are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others.

@@

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%. It should be reiterated that the framework presented
in fig.~\ref{fig:concept} is amenable to changes \emph{e.g.} it is
possible to use distinct trees if working with distinct clades (such as
pollination networks) or, alternatively, the measure of similarity may
not be phylogeny, and can be replaced by information on foraging
(Beckerman et al., 2006), cell-level mechanisms (Boeckaerts et al.,
2021), or a combination of traits and phylogenetic structure (Stock,
2021). Most importantly, although we focus on a trophic system, it is an
established fact that different (non-trophic) interactions do themselves
interact with and influence the outcome of trophic interactions (see
\emph{e.g.} Kawatsu et al., 2021; Kéfi et al., 2012). Future development
of metaweb inference techniques should cover the prediction of multiple
interaction types.

@@

We use data from the European metaweb assembled by Maiorano et al.
(2020a). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
The European metaweb was filtered using the Global Biodiversity
Information Facility (GBIF) taxonomic backbone (GBIF Secretariat, 2021)
so as to contain only terrestrial and semi-aquatic mammals. As all
species had valid matches to the GBIF taxonomy it was used as the
backbone for the remaining reconciliation steps namely, the mammalian
consensus supertree by Upham et al. (2019) (which is used for the
knowledge transfer step) and for the Canadian species list---which was
extracted from the International Union for Conservation of Nature (IUCN)
checklist, and corresponds to the same selection criteria that was
applied by Maiorano et al. (2020a) in the European metaweb.

After taxonomic cleaning and reconciliation the mammalian European
metaweb has 260 species, and the Canadian species pool 163; of these, 17
(about 4\% of the total) are shared, and 89 species from Canada (54\%)
had at least one congeneric species in Europe. The similarity for both
species pools predictably increases with higher taxonomic order, with
19\% of shared genera, 47\% of shared families, and 75\% of shared
orders; for the last point, Canada and Europe each had a single unique
order (\emph{Didelphimorphia} for Canada, \emph{Erinaceomorpha} for
Europe).

@@

The core point of our method is the transfer of knowledge of a known
ecological network, in order to predict interactions between species
from another location for which the network is unknown (or partially
known). In fig.~\ref{fig:concept} the grey text boxes give a high-level
overview of the approach; in the example around which this manuscript is
built The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
(Holm, 2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information than we might think; as such, even algorithms that are
disconnected from a model can make correct guesses most of the time
(Halevy et al., 2009); in fact, in an instance of ecological forecasting
of spatio-temporal systems, model-free approaches (\emph{i.e.} drawing
all of their information from the data) outperformed model-informed ones
(Perretti et al., 2013).

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy et al., 2019), in that the metaweb representation is easily
transferable. Specifically, we use Random Dot Product Graph model
(hereafter RDPG; S. J. Young \& Scheinerman, 2007) to create a number of
latent variables that can be combined into an approximation of the
network adjacency matrix. RDPG is known to capture the evolutionary
backbone of food webs (Dalla Riva \& Stouffer, 2016), resulting in
strong phylogenetic signal in RDPG results; in other words, the latent
variables of an RDPG can be mapped onto a phylogenetic tree, and
phylogenetically similar predators should share phylogenetically similar
preys. In addition, recent advances show that the latent variables
produced this way can be used to predict \emph{de novo} interactions.
Interestingly, the latent variables do not need to be produced by
decomposing the network itself; in a recent contribution, Runghen et al.
(2021) showed that deep artificial neural networks are able to
reconstruct the left and right subspaces of an RDPG, in order to predict
human movement networks from individual/location metadata and opens up
the possibility of using additional metadata as predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD; Halko et al., 2011) on the adjacency matrix.
SVD is an appropriate embedding of ecological networks, which has
recently been shown to both capture their complex, emerging properties
(Strydom, Dalla Riva, et al., 2021) and to allow highly accurate
prediction of the interactions within a single network (Poisot, Ouellet,
et al., 2021). Under SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (with \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.

@@

An RDPG estimates the probability of observing interactions between
nodes (species) as a function of the nodes' latent variables, and is a
way to turn an SVD (which decompose one matrix into three) into two
matrices that can be multiplied to provide an approximation of the
network. The latent variables used for the RDPG, called the left and
right subspaces, are defined as
\(\mathscr{L} = \mathbf{U}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{V}'\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R} = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R} \approx \mathbf{A}\). Using a rank of 1 for the
t-SVD provides a first-order approximation of the network. One advantage
of using an RDPG rather than an SVD is that the number of components to
estimate decreases; notably, one does not have to estimate the singular
values of the SVD. Furthermore, the two subspaces can be directly
multiplied to yield a network.

@@

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
Maiorano et al. (2020a) and O'Connor et al. (2020), this seems like a
reasonable assumption, although one that does not hold for all metawebs.
We used the thresholding approach presented in Poisot, Ouellet, et al.
(2021), and picked a cutoff that maximized Youden's \(J\) statistic (a
measure of the informedness (trust) of predictions; Youden (1950)); the
resulting cutoff was 0.22, and gave an accuracy above 0.99. In Supp.
Mat. 1, we provide several lines of evidence that using the entire
network to estimate the threshold does not lead to overfitting; that
using a subset of species would yield the same threshold; that
decreasing the quality of the original data by adding or removing
interactions would minimally affect the predictive accuracy of RDPG
applied to the European metaweb; and that the networks reconstructed
from artificially modified data are reconstructed with the correct
ecological properties.

@@

Because the confidence intervals on the inferred trait space are
probably over-estimates, we decided to apply a thresholding step to the
interactions after the data inflation (see fig.~\ref{fig:thresholds}
showing the effect of varying the cutoff on \(P(i \rightarrow j)\)).
Cirtwill \& Hambäck (2021) proposed a number of strategies to threshold
probabilistic networks. Their methods assume the underlying data to be
tag-based sequencing, which represents interactions as co-occurrences of
predator and prey within the same tags; this is conceptually identical
to our Bernoulli-trial based reconstruction of a probabilistic network.
We performed a full analysis of the effect of various cutoffs, and as
they either resulted in removing too few interactions, or removing
enough interactions that species started to be disconnected from the
network, we set this threshold for a probability equivalent to 0 to the
largest possible value that still allowed all species to have at least
one interaction with a non-zero probability. The need for this slight
deviation from the Cirtwill \& Hambäck (2021) method highlights the need
for additional development on network thresholding.

@@

@@

One important aspect in which Europe and Canada differ (despite their
comparable bioclimatic conditions) is the degree of the legacy of human
impacts, which have been much longer in Europe. Nenzén et al. (2014)
showed that even at small scales (the Iberian peninsula), mammal food
webs retain the signal of both past climate change and human activity,
even when this human activity was orders of magnitude less important
than it is now. Similarly, Yeakel et al. (2014) showed that changes in
human occupation over several centuries can lead to food web collapse.
Megafauna in particular seems to be very sensitive to human arrival
(Pires et al., 2015). In short, there is well-substantiated support for
the idea that human footprint affects more than the risk of species
extinction (Marco et al., 2018), and can lead to changes in interaction
structure. Yet, owing to the inherent plasticity of interactions, there
have been documented instances of food webs undergoing rapid
collapse/recovery cycles over short periods of time (Pedersen et al.,
2017). The embedding of a network, in a sense, embeds its
macro-evolutionary history, especially as RDPG captures ecological
signal (Dalla Riva \& Stouffer, 2016); at this point, it is important to
recall that a metaweb is intended as a catalogue of all potential
interactions, which should then be filtered (Morales-Castilla et al.,
2015). In practice (and in this instance) the reconstructed metaweb will
predict interactions that are plausible based on the species'
evolutionary history, however some interactions would/would not be
realized due to human impact.

@@

prioritized. While RDPG was able to maintain very high predictive power
when interactions were missing, the addition of false positive

@@

\textbf{Authors' contributions:} TS, SB, and TP designed the study and
performed the analysis; GVDR, MF, and RR provided additional feedback on
the analyses. DC, BM, and FB helped with data collection. All authors
contributed to writing and editing the manuscript.
